Debugging by \textit{lastChange}

### 2010-10-22 papers/pldi2011/sigplanconf-template.tex
some changes for section 4

%---------------------------------------------------------------------------------------------------
\section{JavaScript implementation}
We implemented \textit{Querypointer}, an Extension to Firebug JavaScript debugger, which provides \textit{lastChange} functionality. Firebug is an add-on of Firefox browser. Firefox JavaScript engine provides a JavaScript debugging interface and \textit{Querypointer} is developed over this interface. The debugging support in JavaScript comparing to more traditional languages such as Java and C\# is less evolved. Some part of this flaw is due to the dynamic nature of JavaScript. For example, JavaScript does not have classes and therefore setting watchpoint on properties before program execution is not possible. We explain how we could overcome these limitations and provide required operations for \textit{lastChange} algorithm implementation. Moreover, we breifly describe mechanisms for automated execution reproduction, data collection and presentation.

\subsection{Algorithm technical details}
\subsubsection{\texttt{objectId()} operation}
\texttt{objectId()} operation receives a regular JavaScript object and returns an id which is unique during one execution. This operation keeps the object id in an object property with name \texttt{\_objectId}. Once this operation receives an object, it returns \texttt{\_objectId} value if this property is already defined, otherwise it generates a new id and sets this property object. To set the property, it calls \texttt{defineProperty} standard JavaScript function. \texttt{defineProperty} receives an argument which specifies whether the property is enumerable or not. By setting the enumerable field \texttt{false} for \texttt{\_objectId}, this property will not apear in \texttt{for(property in object)} loops and therfore has no effect on the program execution. Consider that \texttt{\_objectId} is not set for all objects but only those objects that need an id.

\subsubsection{\texttt{scopeId()} operation}
The scope id is kept as a variable with name \texttt{\_scopeId} in the scope. We explained that we set a hook at the first instruction of the function which defines the variable. Once this hook hits, \texttt{\_scopeId} is set by calling \texttt{eval()} function. For example, executing \texttt{eval( var \_scopeId = 10)} creates a variable with name \texttt{\_scopeId} and value \texttt{10} in the local scope which is our interesting scope.

\subsubsection{\texttt{setPropertyChangeHook()} operation}
We implement a slightly different version of this operation. Setting a flag into the Firefox JavaScript engine we can get the place an object created including the JavaScript file url and line number (e.g., myFile.js, line 24). We can assume that an object is created in the same place in the next re-executions. This assumption does not lead to a wrong answer. We will discuss it in the next section when we discuss the effect of non-determinism. Having this assumption, it is enough to set property change hook for the objects created at one place.

The Firefox JavaScript engine supports watching property changes in an object. Every object has a \texttt{watch()} function which receives two parameters, a property name and another function as the callback. Whenever the property with the given name changes the calleback function is called. The hook is set by this function remains enabled even if the property is deleted and defined again. This function can only be applied to avaialbe objects. However, at the beginning of execution, no object excepting the global object is available.

If we could get a reference to the object just after its creation, we could use \texttt{watch()} function to monitor property changes in the object. In the next section we explain the algorithm we devised for this purpose by parsing and analyzing the code. This algorithm could be replaced by a much simpler one if Firefox debugging interface gives us the program counter of object creation instead of line number and also provides object creation hook.
By these assumptions, it was enough to set a hook on the object creation location. Once the execution reaches this location, by enabling object creation hook, we could get a reference to the new object.

%One solution is to assume that jsd has features that simplify it:first: Setting the flag (object track) returns the exact location of object creation instead of the line number.second: It is possible to set hook on object creation and this returns a reference to the created object.
 %Then we can explain it like this: We set a breakpoint at the location object is created and we enable the object creation hook.  We do steppingfor one instruction and we get the reference.

 %I was thinking that perhaps Fireforx 4.0 may support these features.I know that in the new version of JS (1.8.5) and Fireforx 4.0 there is a new function Object(), which is the main object constructor. If for all cases of object creations ({}, new, function),  the engine first calls this function then by modifying or setting a hook in this function we can track object creation. Do you know anyone in Mozilla team who can answer this question?Do you know whether firefox 4 still supports objectTrace flag or not?

@@

\end{verbatim}
\caption{A few cases shows possible object creations.}
\label{fig:objectCreation}
\end{figure}

>>>>

%---------------------------------------------------------------------------------------------------
\section{JavaScript implementation}
To verify the \textit{lastChange} algorithm we
implemented it in an extension to the Firebug
JavaScript debugger\cite{firebug-version-1.6}. 
Firebug itself is an extension of the Firefox browser\cite{firefox-version-3.6}. The Firefox
JavaScript engine provides a JavaScript debugging interface and
\textit{Querypointer} is developed over this interface.

\subsection{Algorithm technical details}
\subsubsection{\texttt{objectId()} operation}
\texttt{objectId(obj)} first checks the argument \texttt{obj} for a property \texttt{\_objectId}.
It returns the value if this property is already defined,
otherwise it generates a new id and sets this property. The value of the id is simply an integer incremented 
for each new \texttt{\_objectId} needed. \textit{CHECK THAT} To set
the property, we call the \texttt{defineProperty} standard JavaScript
function. \texttt{defineProperty} receives an argument which specifies
whether the property is enumerable or not. By setting the enumerable
field \texttt{false} for \texttt{\_objectId}, this property will not
apear in \texttt{for(property in object)} loops and therfore has no
effect on the program execution. Note that \texttt{\_objectId} is
not set for all objects but only those objects that need an id.

\subsubsection{\texttt{scopeId()} operation}
The scope id is kept as a variable with name \texttt{\_scopeId} in the
scope. We explained that we set a hook at the first instruction of the
function which defines the variable. Once this hook hits,
\texttt{\_scopeId} is set by calling JavaScript's dynamic compilation function \texttt{eval()}. For
example, executing \texttt{eval( var \_scopeId = 10)} creates a
variable with name \texttt{\_scopeId} and value \texttt{10} in the
scope of the \texttt{eval()} call, which is our interesting scope.

\subsubsection{\texttt{setPropertyChangeHook()} operation}
The Firefox JavaScript engine supports watching property changes in an
object. Every object has a function \texttt{watch(propName, callback)}
which receives two parameters, a property name and function.  Whenever
the property with the given name changes, the \texttt{callback} function is
called. The hook set by this function remains enabled even if the
property is deleted and defined again.

For our purposes, the \texttt{watch()} function only covers the case
of global object properties. At the beginning of execution, no object
excepting the global object is available. Therefore we defined and
implemented \texttt{setPropertyChangeHook()} to work in the rest of
the cases.

The Firefox JavaScript engine does not current support
\texttt{setPropertyChangeHook()} (see however \cite{bugzilla}).  For
the our \textit{lastChange} prototype we created a version of
\texttt{setPropertyChangeHook()} using only features of the current
engine.  The basic strategy is to get a reference to the object just
after its creation, then use \texttt{watch()} function to monitor
property changes in the object.  Setting a flag into the Firefox
JavaScript engine we can get file URL and line number for each object
creation (e.g., myFile.js, line 24). We set a breakpoint on this line
and analyze the source code to determine which object was created.

@@

\end{verbatim}
\caption{Examples of three types of JavaScript object creatin statements.}
\label{fig:objectCreation}
\end{figure}
